---
title: "<b>Portugal Market Analysis</b>"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    orientation: columns
    vertical_layout: fill
    theme: simplex
editor_options: 
  chunk_output_type: console
runtime: shiny
---
<!-- theme alternatives: "cerulean",  "journal", "readable", "simplex", "united", "yeti" -->

<style>
  body {
    text-align: justify;
    font-size: 15px;
  }
  h2 {
    font-size: 18px;
  }
</style>

```{r SETUP, message=FALSE, include=FALSE}
#GENERAL CODE CHUNK SETUP
knitr::opts_chunk$set(echo = FALSE, message=FALSE)

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#PACKAGE INSTALLATION AND LOADING
#-the essential packages required to load the document are markdown, knitr, flexdashboard and pacman.
#-the following code checks for their installation, installing them if required, and proceeds
#-to install/load other necessary packages.

if(!require("markdown")) install.packages("markdown")
if(!require("knitr")) install.packages("knitr")
if(!require("flexdashboard")) install.packages("flexdashboard")
if(!require("pacman")) install.packages("pacman") 
#--pacman function essential for loading and installing (if not already installed) packages together

pacman::p_load(
  knitr,           #--essential for rmarkdown document modification
  tidyverse,       #--contains ggplot2, readr, tibble, dplyr, magrittr
  reactable,       #--for creation of large tables
  janitor,         #--manipulation, summarisation of messy data, plus duplicate detection
  inspectdf,       #--dataframe inspection, relative to value (used in dataset analysis)
  highcharter,     #--interactive data visualisation
  htmltools,       #--website page options (zoom, scrolling)
  collapsibleTree, #--for happiness factor collapsible tree
  broom,           #--needed to convert linear model for interpolation line into dataframe
  dygraphs,        #--package for time series model creation
  xts,             #--data format required by dygraphs
  shiny,           #--"runtime: shiny" in JMAL header already loads the library, useful for debugging w/o                       knitting
  shinyWidgets,    #--additional customisation for shiny elements
  readxl,
  viridis
  )
```


```{r FUNCTIONS & DATA IMPORTING, message=FALSE, include=FALSE}
#FUNCTIONS
#-".csv" extension remover, for variable list
ext_remove <- 
  function(x){
    substr(x, 1, nchar(x)-4)
  }

xls_remove <-
  function(x){
    substr(x, 1, nchar(x)-5)
  }

#-----#

#HIGHCHARTER LINE CHART FORMAT
lnplot_custom <- function(x){
  x %>% 
    hc_title(style=list(
               fontWeight="bold",
               fontSize= "15px",
               fontFamily= "Lato"
              )
             ) %>% 
    hc_size(width = NULL, height = 350) %>% 
    hc_yAxis(
      type="linear",
      lineWidth=5,
      labels=list(format="{value}"),
      lineColor="#ecf0f1") %>%
    hc_xAxis(
      type="linear",
      lineWidth=5,
      labels=list(format="{value}"), 
      lineColor="#ecf0f1") %>% 
    hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>
          Year: {point.x:.f}<br>Population: {point.y:.f}") %>% 
    hc_plotOptions(.,line=list(color="#db240c")) 
}

#-----#

#HIGHCHARTER() BAR PLOT FORMAT
barplot_custom <- function(x){
  x %>% 
    hc_title(style=list(
               fontWeight="bold",
               fontSize= "15px",
               fontFamily= "Lato"
              )
             ) %>% 
    hc_size(width = NULL, height = 450) %>% 
    hc_yAxis(
      type="linear",
      lineWidth=5,
      labels=list(format="{value}"),
      lineColor="#ecf0f1") %>%
    hc_xAxis(
      type="linear",
      lineWidth=5,
      labels=list(format="{value}"), 
      lineColor="#ecf0f1") %>%
    hc_tooltip(shared = TRUE, crosshairs = TRUE, sort = TRUE) 
}
  
#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#

#DATA IMPORTING
#when downloading the file, the working directory should automatically set to the path where the Markdown file is saved. if that does not occur, copy the Markdown from the dropdown menu between the "" and run the single line of code.
setwd("C:/Users/busop/All Files - ORIGINAL FILES OF DESKTOP LINKS/Github/market-analysis-2024-2025/Rmarkdown")

dsets <- list.files(path="./Datasets", pattern = "*.csv", full.names = FALSE)   
#vector with file names

dsets <-
  #assign the value contained in each dataset to its respective name
  setNames(
    lapply(list.files(path= "./Datasets", pattern="*.csv", full.names = TRUE), read.csv), 
    #reading each csv file, same code as dsets w/ exception that it takes full file path rather than just the name
    lapply(dsets, ext_remove)
    #removing ".csv" file extension to create variable list, function above
  )

#same concept reapplied for xlsx files, useful to separate the two
xlsx_dsets <- list.files(path="./Datasets", pattern = "*.xlsx", full.names = TRUE)

xlsx_dsets <-
  setNames(
    lapply(xlsx_dsets, read_xlsx),
    lapply(list.files(path="./Datasets", pattern = "*.xlsx", full.names = FALSE), xls_remove)
  )

#final dsets vectors contains variables associated w/ respective dataset:
#-global edits done to reduce variable names, since discarded data not required.
#-method flexible to increases, decreases or variations in order to dataset directory.
```



<!-- INTRODUCTION PAGE -->

# Introduction

<!-- DEFINITIONS SECTION -->

## Definitions and Graph {data-width="450"}

### **Research Question and Definitions**

Fruit Snacks, Snack Bars, Sweet Biscuits in Portugal and the strategy of *Vieira de Castro & Filhos Lda.* in that market.

- What is a snack?
  - "A small meal or amount of food, usually eaten in a hurry." (Oxford Dictionary)
- Types of snack considered:
  - Fruit Snacks (dried fruit, processed fruit snacks).
  - Snack Bars (cereal bars, protein/energy bars, fruit and nut bars, other snack
  bars).
  - Sweet Biscuits (chocolate coated biscuits, cookies, filled biscuits, plain
  biscuits, wafers).
- Outline of the snack market:
  - The market considered encloses the three categories above, including all their subcategories.
  - The market is valued according to retail selling price (RSP) and includes any
  applicable taxes (definition from Euromonitor).
  - Statistical indicators considered pertain to the subset specified, at most representing the general food    manufacture sector. 

<!-- INDEX SECTION -->

### **Index**
```{r REPORT FACTOR WEB}
Index <- 
  data.frame(
    parent = c(rep("Socio-Demographic Analysis", 4), rep("Economic Overview", 3), rep("Supply", 3),
               rep("Demand", 4), rep("Vieira de Castro & Filhos Lda.", 2)),
    child = c("Demographic Data", "General Health","Portuguese Economy", "Parallel with Italy and Spain",
              "Real GDP and RGDP Growth Rate", "Unemployment Rate","Inflation Rate",
              "Until 2012", "From 2016 to 2018", "2022 - COVID19 Effect",
              "Snack Categories", "Category Retail Revenue", "Price-Volume Variations", "Income-Volume Variations",
              "A", "B"
              )
  )

collapsibleTree(Index, hierarchy = c("parent", "child"), collapsed = FALSE, 
                width = 600, height = NULL, tooltip = FALSE, fontSize= 10, fill= "#db240c")

```

<!-- ANALYSIS METHODS SECTION -->

## Analysis Methods {data-width="550"}

<!-- PREMISES TO INFERENCE SUBSECTION -->

### **Premises to Inference**

In order to achieve reliable conclusions, a rigid information selection process was adhered to, in accordance with the following principles:

- Pertaining to quantitative information and assessing qualitative sources with
care.
- Thoroughly checking the validity, completeness and accessibility of the
information gathered online.
- Ensuring the reliability of the external, secondary sources used.
  - Information mainly gathered from Eurostat, Passport and OurWorldinData.
- Selecting truthful graphical representationos, avoiding indirect manipulation of data.
- Use of standardised values.

<!-- CHALLENGES FACED SUBSECTION -->

### **Challenges Faced**

Throughout the analysis, a series of problems were incurred that complicated 
operations. It is to be expected that a research on a topic as expansive as a
market leads to fragmented conclusions or unsatisfactory outcomes. Some issues
faced were:

- Lack of public data for the company of study.
- Discordant data between sources.
- Lack of confidence in some data correlation.

<!-- OVERCOMING CHALLENGES SUBSECTION -->

### **Overcoming Challenges**

To face the set of complications mentioned previously, the initial approach was 
to evaluate the quality of our available sources. 

- Creation of a hierarchy of source quality, peaking with governmental-linked information, followed up by       trustworthy data providers.
- Contacting data providers (e.g. Vieira itself) to obtain relevant information.
  - Resorting to a brand analysis for lack of response and of alternatives.
- Cross-referencing information sources.

<!-- SOCIO-DEMOGRAPHICS PAGE -->

# Socio-Demographic Factors

<!-- POPULATION SECTION -->

## Population {.tabset .tabset-fade data-width="650"}

<!-- TAB 1 DEMOGRAPHICS -->

### **Demographic Data**

**Population Growth**:

- Exponential demographic increase in Portugal during the 20th century.
- Influenced by improved living conditions in Europe after the Industrial Revolutions, including economic,      social, political, educational, and sanitary advancements.
- Post-WWII period of peace contributed to stability, fostering demographic growth.

**Population Density Trends**:

- Fluctuating trends observed over 200 years.
- Noticeable shift in the 1960s linked to mass emigration during the Estado Novo dictatorship.

**Immigration Dynamics**:

- Highlighted the significance of foreigner influx into Portugal.
- Trend variation after the 2008 global financial crisis due to more restrictive immigration policies.

<div style="height: 50px;"></div>

```{r PORTUGAL DEMOGRAPHICS}
#CREATION OF CLEAN POPULATION, POPULATION DENSITY, LIFE EXPECTANCY, IMMIGRATION STOCK DATASETS
clean_population <- 
  dsets[["portugal_population"]] %>% 
  filter(Year >= 1800 & Entity == "Portugal") %>% 
  select(-Code, -Entity) %>% 
  rename(`Population` = Population..historical.) %>% 
  hchart("line", hcaes(`Year`, `Population`)) %>% 
  hc_title(text = "Portuguese Population") %>% 
  lnplot_custom(.) %>% 
  hc_yAxis(title=list(text="<b>Population [inhab]</b>", align="middle"),
           labels = list(
                      formatter = JS("function() {
                          if (this.value >= 1000000) {
                            return (this.value / 1000000) + ' mil';
                          } else if (this.value >= 1000) {
                            return (this.value / 1000) + ' k';
                          }
                          return this.value;
                        }")
                      )
          ) %>% 
  hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>Population: {point.y:.f} inhab") %>% 
  hc_caption(
    text = "<b>Population</b> = number of inhabitants of a country, at any given year.",
    style = list(marginTop = "0px")
  )
  
  
clean_population_density <-
  dsets[["portugal_population_density"]] %>% 
  filter(Year >= 1800 & Entity == "Portugal" & Year <= 2023) %>% 
  select(-Code, -Entity) %>% 
  rename(`Population Density` = Population.density) %>% 
  hchart("line", hcaes(`Year`, `Population Density`)) %>% 
  hc_title(text = "Portuguese Population Density") %>% 
  lnplot_custom(.) %>% 
  hc_yAxis(title = list(text = "<b>Population Density [inhab/kmsq]</b>", align = "middle")) %>% 
  hc_xAxis(title = list(text = "<b>Year</b>", align = "middle")) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>Population Density: {point.y:.1f} inhab/kmsq") %>% 
  hc_caption(
    text = "<b>Population Density</b> = population per square kilometer by country.",
    style = list(marginTop = "0px")
  )

clean_immigration <-
  dsets[["portugal_migrant_stock"]]%>% 
  filter(Year >= 1800 & Entity == "Portugal" & Year <= 2023) %>% 
  select(-Code, -Entity) %>% 
  rename(`International Immigrants` = Total.number.of.international.immigrants) %>% 
  hchart("line", hcaes(`Year`, `International Immigrants`)) %>% 
  hc_title(text = "Portuguese International Immigrants") %>% 
  lnplot_custom(.) %>% 
  hc_yAxis(title = list(text = "<b>N° Immigrants [inhab]</b>", align = "middle"),
           labels = list(
                      formatter = JS("function() {
                          if (this.value >= 1000000) {
                            return (this.value / 1000000) + ' mil';
                          } else if (this.value >= 1000) {
                            return (this.value / 1000) + ' k';
                          }
                          return this.value;
                        }")
                      )
      ) %>% 
  hc_xAxis(title = list(text = "<b>Year</b>", align = "middle")) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>Immigrant Stock: {point.y:.f} inhab") %>% 
     hc_caption(
    text = "<b>International Immigration Stock</b> = people living in a given country who were born in another country, recorded at mid-year by age and sex and by region, country or area of destination.",
    style = list(marginTop = "0px")
  )

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC
pop_graph_choice <- c('Population', 'Population Density', 'International Immigrants')
pop_graph_code <- list(clean_population, clean_population_density, clean_immigration)

pop_graph_ui <-
  fluidPage({
    sidebarLayout(
      sidebarPanel(
        width = 4,
        pickerInput(
          inputId = "pop_selection",
          label = "Demographics Graphs",
          choices = pop_graph_choice,
          selected = pop_graph_choice[1],
          options = list(style = "btn-danger"),
          width = "100%"
        )
      ),
      
      mainPanel(
        width = 8,
        highchartOutput(outputId = "pop_highchart")
      ),
      
      position = c("left", "right")
    )
  })

pop_graph_server <- function(input, output){
  pop_selected <- reactive({
    pop_graph_code[pop_graph_choice %in% input$pop_selection]
  })
  
  output$pop_highchart <-
    renderHighchart({
      temp <- pop_selected() 
      temp[[1]]
    })
  }
      
shinyApp(ui = pop_graph_ui, server = pop_graph_server)
```

<!-- TAB 2 HEALTH -->

### **General Health**
**Life Expectancy**:

- Dramatic increase post-WWII, marked by gradual, quasi-constant growth with reduced volatility over time.
- Improvements linked to health system regulations, vaccination campaigns, and better living conditions.

**Caloric Supply**:

- Average daily caloric supply mostly increased but diverged sharply in the late 1970s–1980s.
- This fluctuation likely due to economic/social disruptions, including the 1970s inflationary crisis and       immigration from former African colonies.

**Obesity Trends**:

- Obesity in Portugal shows a steady, linear growth.
- Despite fluctuations in caloric supply, obesity growth suggests increasing availability and consumption of    calorie-dense, economical foods, such as snacks.

<div style="height: 50px;"></div>

```{r PORTUGAL HEALTH}
clean_life_expectancy <-
  dsets[["portugal_life_expectancy"]] %>% 
  filter(Year >= 1800 & Entity == "Portugal" & Year <= 2023) %>% 
  select(-Code, -Entity) %>% 
  rename(`Life Expectancy` = Period.life.expectancy.at.birth...Sex..all...Age..0) %>% 
  hchart("line", hcaes(`Year`, `Life Expectancy`)) %>% 
  hc_title(text = "Portuguese Life Expectancy") %>% 
  lnplot_custom(.) %>% 
  hc_yAxis(title = list(text = "<b>Life Expectancy  [yrs]</b>", align = "middle")) %>% 
  hc_xAxis(title = list(text = "<b>Year</b>", align = "middle")) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>Life Expectancy: {point.y:.1f} years") %>% 
   hc_caption(
    text = "<b>Life Expectancy</b> = estimated time period an average individual (normalised by sex and age) is expected to live.",
    style = list(marginTop = "0px")
  )

clean_calorie_supply <-
  dsets[["portugal_daily_caloric_supply"]] %>% 
  filter(Year >= 1800 & Entity == "Portugal" & Year <= 2023) %>% 
  select(-Code, -Entity) %>% 
  rename(`Daily Calorie Supply` = Daily.calorie.supply.per.person) %>% 
  hchart("line", hcaes(`Year`, `Daily Calorie Supply`)) %>% 
  hc_title(text = "Portuguese Daily Calorie Supply") %>% 
  hc_yAxis(title = list(text = "<b>Average Daily Calorie Supply [kcal]</b>", align = "middle")) %>% 
  hc_xAxis(title = list(text = "<b>Year</b>", align = "middle")) %>% 
  lnplot_custom(.) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>Average Daily Calorie Supply: {point.y:.1f} kcal") %>% 
  hc_caption(
    text = "<b>Per Capita Daily Calorie Supply</b> = amount of calories available to an average person, not   necessarily the calories actually consumed by that person.",
    style = list(marginTop = "0px")
  )
  
clean_obesity_rate <-
  dsets[["portugal_overweight_adults"]]%>% 
  filter(Year >= 1800 & Entity == "Portugal" & Year <= 2023) %>% 
  select(-Code, -Entity) %>% 
  rename(`Overweight Share of Adults` =
           Prevalence.of.obesity.among.adults..BMI....30..crude.estimate........Sex..both.sexes...Age.group..18...years) %>% 
  hchart("line", hcaes(`Year`, `Overweight Share of Adults`)) %>% 
  hc_title(text = "Portuguese Obesity Rate") %>% 
  hc_yAxis(title = list(text = "<b>Obesity Rate [%]</b>", align = "middle")) %>% 
  hc_xAxis(title = list(text = "<b>Year</b>", align = "middle")) %>% 
  lnplot_custom(.) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>Obesity Rate: {point.y:.1f}%") %>% 
  hc_caption(
    text = "<b>Obesity Rate</b> = percentage of adults aged 18+ years with a body mass index (BMI) of 30 kg/m2 or higher.",
    style = list(marginTop = "0px")
  )
  
#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC
health_graph_choice <- c('Life Expectancy', 'Daily Calorie Supply', 'Obesity Rate')
health_graph_code <- list(clean_life_expectancy, clean_calorie_supply, clean_obesity_rate)

health_graph_ui <-
  fluidPage({
    sidebarLayout(
      sidebarPanel(
        width = 4,
        pickerInput(
          inputId = "health_selection",
          label = "Health Infographics",
          choices = health_graph_choice,
          selected = health_graph_choice[1],
          options = list(style = "btn-danger"),
          width = "100%"
        )
      ),
      
      mainPanel(
        width = 8,
        highchartOutput(outputId = "health_highchart")
      ),
      
      position = c("left", "right")
    )
  })

health_graph_server <- function(input, output){
  health_selected <- reactive({
    health_graph_code[health_graph_choice %in% input$health_selection]
  })
  
  output$health_highchart <-
    renderHighchart({
      temp <- health_selected() 
      temp[[1]]
    })
  }
      
shinyApp(ui = health_graph_ui, server = health_graph_server)
```

<!-- TAB 3 WEALTH -->

### **Portuguese Economy**

**GDP Per Capita**:

- Indicator of average wealth per individual, though it does not account for wealth distribution.
- Shows a general upward trend, with deviations during global crises (e.g., 2008 banking crisis, 2020 COVID-19   pandemic).
- Reflects improving living conditions and greater attainable wealth for individuals.

**Human Development Index (HDI**:

- Introduced to address GDP per capita's limitations in capturing broader social factors.
- Confirms the trend of gradual growth in living standards, with minimal time-based variation.

**Poverty Statistics**:

- Recent data highlights the percentage of the population below a fixed poverty line (**€7,095 annually or      €591 monthly**).
- Indicates a reduction in poverty post-Euro introduction, linked to Portugal's alignment with EU economic      regulations. 

<div style="height: 50px;"></div>

```{r PORTUGAL WEALTH}
#CLEAN GDP PRO CAPITA, HDI DATASETS
clean_gdp_capita <-
  dsets[["portugal_gdp_per_capita"]] %>% 
  filter(Year >= 1800 & Entity == "Portugal" & Year <= 2023) %>% 
  select(-Code, -Entity) %>% 
  rename(`GDP per capita` = GDP.per.capita..PPP..constant.2017.international...) %>% 
  hchart("line", hcaes(`Year`, `GDP per capita`)) %>% 
  lnplot_custom(.) %>% 
  hc_title(text = "Portuguese GDP per Capita") %>% 
  hc_yAxis(title = list(text = "<b>GDP per Capita [int $/inhab]</b>", align = "middle"),
           labels = list(
                      formatter = JS("function() {
                          if (this.value >= 1000000) {
                            return (this.value / 1000000) + ' mil';
                          } else if (this.value >= 1000) {
                            return (this.value / 1000) + ' k';
                          }
                          return this.value;
                        }")
                      )
           ) %>% 
  hc_xAxis(title = list(text = "<b>Year</b>", align = "middle")) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>GDP per capita: {point.y:.f} $/inhab") %>% 
  hc_caption(
    text = "<b>GDP per Capita</b> = gross domestic product converted to international dollars using purchasing power parity rates, put into relation with population size.",
    style = list(marginTop = "0px")
  )

clean_hdi <-
  dsets[["portugal_hdi"]] %>% 
  filter(Year >= 1800 & Entity == "Portugal" & Year <= 2023) %>% 
  select(-Code, -Entity) %>% 
  rename(`HDI` = Human.Development.Index) %>% 
  hchart("line", hcaes(`Year`, `HDI`)) %>% 
  hc_title(text = "Portuguese Human Development Index") %>% 
  hc_yAxis(title = list(text = "<b>Human Development Index</b>", align = "middle")) %>% 
  hc_xAxis(title = list(text = "<b>Year</b>", align = "middle")) %>% 
  lnplot_custom(.) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>HDI: {point.y:.f}") %>% 
  hc_caption(
    text = "<b>Human Development Index</b> = summary measure of key dimensions of human development (mean years of schooling, expected years of schooling, life expectancy at birth and gross national income per capita).",
    style = list(marginTop = "0px")
  )

clean_poverty_percentage <-
  dsets[["portugal_population_poverty"]] %>% 
  filter(Year >= 1800 & Entity == "Portugal" & Year <= 2023) %>% 
  select(-Code, -Entity) %>% 
  rename(`Poverty Ratio` = Poverty.headcount.ratio.at.national.poverty.lines....of.population.) %>% 
  hchart("line", hcaes(`Year`, `Poverty Ratio`)) %>% 
  hc_title(text = "Portuguese Relative Population Poverty Percentage [%]") %>% 
  hc_yAxis(title = list(text = "<b>National Poverty Headcount Ratio [%]</b>", align = "middle")) %>% 
  hc_xAxis(title = list(text = "<b>Year</b>", align = "middle")) %>% 
  lnplot_custom(.) %>% 
  hc_tooltip(., 
       headerFormat = "",
       pointFormat = 
         "<b>{point.name}</b><br>Year: {point.x:.f}<br>Poverty Percentage: {point.y:.1f}%") %>% 
  hc_caption(
    text = "<b>National poverty headcount ratio</b> =  percentage of the population living below the national poverty line(s) [591 € monthly income]. <br>National estimates are based on population-weighted subgroup estimates from household surveys.",
    style = list(marginTop = "0px")
  )

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC
wealth_graph_choice <- c('GDP per Capita', 'Human Development Index', 'National Poverty Headcount Ratio')
wealth_graph_code <- list(clean_gdp_capita, clean_hdi, clean_poverty_percentage)

wealth_graph_ui <-
  fluidPage({
    sidebarLayout(
      sidebarPanel(
        width = 4,
        pickerInput(
          inputId = "wealth_selection",
          label = "Economic Indicators",
          choices = wealth_graph_choice,
          selected = wealth_graph_choice[1],
          options = list(style = "btn-danger"),
          width = "100%"
        )
      ),

      mainPanel(
        width = 8,
        highchartOutput(outputId = "wealth_highchart")
      ),

      position = c("left", "right")
    )
  })

wealth_graph_server <- function(input, output){
  wealth_selected <- reactive({
    wealth_graph_code[wealth_graph_choice %in% input$wealth_selection]
  })

  output$wealth_highchart <-
    renderHighchart({
      temp <- wealth_selected()
      temp[[1]]
    })
  }

shinyApp(ui = wealth_graph_ui, server = wealth_graph_server)
```

<!-- TAB 4 INTERPRETATION -->

### **Interpretation**

**Interpretation**:

- Integrates demographic, sanitary, and economic data with historical and cultural context.
- Divides Portuguese development into three key time periods: pre-1970, 1970–2014, and post-2014.

**Pre-1970**:

- Economy dominated by agriculture and low industrialization.
- Lack of a developed financial sector to support investments.
- Characterized by underdevelopment and minimal structural change.

**1970–2014**:

- Marked by the Carnation Revolution (1974), transitioning Portugal from dictatorship (Estado Novo) to          democracy.
- Structural reforms included agrarian reform and EU integration.
- EU membership enabled access to the single market, boosting exports, increasing capital, and fostering        gradual economic convergence with more advanced EU states.

**Post-2014**:

- Period of economic and social reforms consolidating Portugal’s position within the EU.
- Expansion of tourism and growing demand for local goods.
- Significant financial investments by the government and corporate sectors to access new markets.

**Underlying Developments Across Periods**:

- Steady technological, social, and health improvements, including:
- Careful immigration policies.
- Advances in healthcare technology and services.
- Establishment of a formal education system (Law 46/1986).

<!-- PARALLELS SECTION -->

## Parallels{data-width = "450"}

### **Parallels with Italy and Spain** 



```{r PORTUGAL SPAIN ITALY PARALLELS}
#CLEAN PARALLELS DATASETS
pickerIp_choice <- c("Population Growth Rate", "Life Expectancy", "GDP per Capita")
radioBn_choice <- c("Portugal", "Italy", "Spain")

parallels_population <-
  dsets[["portugal_population_growth_rate"]] %>%
  filter(Year >= 1950 & Entity %in% radioBn_choice) %>%
  select(-Code) %>%
  rename(`Population` = Population.growth.rate)

parallels_population$Entity <-
  factor(parallels_population$Entity, levels = c("Portugal", "Italy", "Spain"))

parallels_population <-
  parallels_population %>% 
  group_by(Entity)

parallels_life_expect <-
  dsets[["portugal_life_expectancy"]] %>%
  filter(Year >= 1950 & Entity %in% radioBn_choice) %>%
  select(-Code) %>%
  rename(`Life Expectancy` = Period.life.expectancy.at.birth...Sex..all...Age..0)

parallels_life_expect$Entity <-
  factor(parallels_life_expect$Entity, levels = c("Portugal", "Italy", "Spain"))

parallels_life_expect <-
  parallels_life_expect %>% 
  group_by(Entity)

parallels_gdp <-
  dsets[["portugal_gdp_per_capita"]] %>%
  filter(Year >= 1800 & Entity %in% radioBn_choice) %>%
  select(-Code) %>%
  rename(`GDP per capita` = GDP.per.capita..PPP..constant.2017.international...)

parallels_gdp$Entity <-
  factor(parallels_gdp$Entity, levels = c("Portugal", "Italy", "Spain"))

parallels_gdp <-
  parallels_gdp %>% 
  group_by(Entity)
  
parallels_graph_code <- list(parallels_population, parallels_life_expect, parallels_gdp)

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC

parallels_graph_ui <-
  fluidPage({
    sidebarLayout(
      sidebarPanel(
        width = NULL,
        pickerInput(
          inputId = "parallels_selection",
          label = "Comparison Factor",
          choices = pickerIp_choice,
          selected = pickerIp_choice[1],
          options = list(style = "btn-danger"),
          width = "100%"
        ),
        checkboxGroupInput(inputId = "radioB_specifics",
                     label = "Countries Compared",
                     choices = radioBn_choice,
                     selected = radioBn_choice[1])
      ),
      
      mainPanel(
        width = NULL,
        highchartOutput(outputId = "parallels_highchart")
      ),
      
      position = c("left", "right")
    )
  })

parallels_graph_server <- function(input, output){
  parallels_topic_choice <- reactive({
    if(input$parallels_selection == pickerIp_choice[1]){
      parallels_graph_code[1]
    } else if(input$parallels_selection == pickerIp_choice[2]){
      parallels_graph_code[2]
    }else{
      parallels_graph_code[(length(parallels_graph_code))]
    }
  })
  
  parallels_scale <- reactive({
    if(input$parallels_selection == pickerIp_choice[1]){
      c(-2, 2)
    }else if(input$parallels_selection == pickerIp_choice[2]){
      c(57, 85)
    }else{
      c(15000, 50000)
    }
  })
  
  parallels_filtered_data <-
    reactive({
      req(input$radioB_specifics)
      temp <- parallels_topic_choice()
      temp <- temp[[1]] %>%
        filter(Entity %in% as.character(input$radioB_specifics))
      scale <- parallels_scale()
      
      if(input$parallels_selection == pickerIp_choice[1]){
        temp %>%
          hchart("line", hcaes(`Year`, `Population`, group = Entity)) %>%
          hc_title(text = "Population Growth Rate Comparison",
                   style=list(
                              fontWeight="bold",
                              fontSize= "15px",
                              fontFamily= "Lato"
                             )
              ) %>%
          hc_yAxis(title = list(text = "<b>Population Growth Rate [%]</b>", align = "middle"),
                   type="linear",
                   lineWidth=5,
                   labels=list(format="{value}%"),
                   lineColor="#ecf0f1") %>%
          hc_xAxis(title = list(text = "<b>Year</b>", align = "middle"),
                   type="linear",
                   lineWidth=5,
                   labels=list(format="{value}"),
                   lineColor="#ecf0f1") %>%
          hc_yAxis(min = scale[1], max = scale[2]) %>% 
          hc_tooltip(.,
                     headerFormat = "",
                     pointFormat =
                       "<b>{point.name}</b><br>Year: {point.x:.f}<br>Growth Rate: {point.y:.2f}%") %>%
          hc_plotOptions(
            line = list(zoneAxis = "x", 
                        zones = list(list(value = 2023),
                                     list(dashStyle = "Dash")
                        )
            )
          ) %>% 
          hc_colors(., c("#db240c", "#046A38", "#FFE900", "#002D72"))
      } else if(input$parallels_selection == pickerIp_choice[2]){
        temp %>%
          hchart("line", hcaes(`Year`, `Life Expectancy`, group = Entity)) %>%
          hc_title(text = "Life Expectancy Comparison",
                   style=list(
                              fontWeight="bold",
                              fontSize= "15px",
                              fontFamily= "Lato"
                             )
                   ) %>%
          hc_yAxis(title = list(text = "<b>Life Expectancy [yrs]</b>", align = "middle"),
                   type="linear",
                   lineWidth=5,
                   labels=list(format="{value} yrs"),
                   lineColor="#ecf0f1") %>%
          hc_xAxis(title = list(text = "<b>Year</b>", align = "middle"),
                   type="linear",
                   lineWidth=5,
                   labels=list(format="{value}"),
                   lineColor="#ecf0f1") %>%
          #hc_size(width = NULL, height = 470) %>% 
          hc_yAxis(min = scale[1], max = scale[2]) %>% 
          hc_tooltip(.,
                     headerFormat = "",
                     pointFormat =
                       "<b>{point.name}</b><br>Year: {point.x:.f}<br>Life Expectancy: {point.y:.1f} years"
                     ) %>%
          hc_colors(., c("#db240c", "#046A38", "#FFE900", "#002D72"))}
      else {
        temp %>%
          hchart("line", hcaes(`Year`, `GDP per capita`, group = Entity)) %>%
          hc_title(text = "GDP per Capita Comparison",
                   style=list(
                              fontWeight="bold",
                              fontSize= "15px",
                              fontFamily= "Lato"
                             )
                   ) %>%
          hc_yAxis(title = list(text = "<b>GDP per Capita [int $/inhab]</b>", align = "middle"),
                   type="linear",
                   lineWidth=5,
                   labels=list(formatter = JS("function() {
                                              if (this.value >= 0) {
                                                return (this.value/1000) + ' k';
                                              }
                                            }")
                              ),
                   lineColor="#ecf0f1") %>%
          hc_xAxis(title = list(text = "<b>Year</b>", align = "middle"),
                   type="linear",
                   lineWidth=5,
                   labels=list(format="{value}"),
                   lineColor="#ecf0f1") %>%
          hc_yAxis(min = scale[1], max = scale[2]) %>% 
          hc_tooltip(.,
                     headerFormat = "",
                     pointFormat =
                       "<b>{point.name}</b><br>Year: {point.x:.f}<br>GDP per Capita: {point.y:.0f} $/inhab"
                     ) %>%
          hc_colors(., c("#db240c", "#046A38", "#FFE900", "#002D72"))
      }
    })
  
  output$parallels_highchart <- renderHighchart({
    req(input$radioB_specifics)
    parallels_filtered_data()
  })
  
}

shinyApp(ui = parallels_graph_ui, server = parallels_graph_server)
```

<!-- ECONOMIC PAGE -->

# Economic Overview

<!-- GRAPHS -->

## General Economic Overview - Graphs {data-width="450"}

###

```{r ECONOMIC OVERVIEW}
#CLEAN ECONOMICS DATASETS
economic_factor_choice <- 
  c("Real GDP", "Real GDP Growth Rate", "Unemployment Rate", "Inflation Rate")

economics_rgdp <-
  dsets[["portugal_real_gdp"]] %>% 
  rename(`Real GDP` = CLVMNACSCAB1GQPT, `Year` = DATE)

economics_rgdp_g_rate <-
  economics_rgdp %>% 
  mutate(`Real GDP Growth Rate` = ((`Real GDP` - lag(`Real GDP`)) / lag(`Real GDP`) * 100)) %>% 
  select(-`Real GDP`)

economics_unemployment <-
  dsets[["portugal_unemployment_rate"]] %>% 
  select(-Code) %>%
  filter(Entity == "Portugal") %>% 
  rename(`Unemployment Rate` = Unemployment..total....of.total.labor.force...modeled.ILO.estimate.)

economics_inflation <-
  dsets[["portugal_inflation"]] %>%
  select(-Code) %>%
  filter(Entity == "Portugal") %>%
  rename(`Inflation Rate`= Inflation..consumer.prices..annual...)

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC
economics_graph_code <- list(economics_rgdp, economics_rgdp_g_rate, economics_unemployment, economics_inflation)

economics_graph_ui <-
  fluidPage({
    sidebarLayout(
      sidebarPanel(
        width = 4,
        pickerInput(
          inputId = "economics_selection",
          label = "General Economic Indicators",
          choices = economic_factor_choice,
          selected = economic_factor_choice[1],
          options = list(style = "btn-danger"),
          width = "100%"
        ),
        chooseSliderSkin("Flat", color = '#db240c'),
        sliderInput(inputId = "economics_year_slider",
                    label = "Year Interval",
                    min = 1990,
                    max = 2023,
                    value = c(1995, 2023),
                    step = 1,
                    sep = "",
                    dragRange = TRUE
                    )
      ),

      mainPanel(
        width = 8,
        highchartOutput(outputId = "economics_highchart")
      ),

      position = c("left", "right")
    )
  })

economics_graph_server <- function(input, output){
  economics_selected_graph <- reactive({
    economics_graph_code[economic_factor_choice %in% input$economics_selection]
  })
  
  economics_filtered_graph <- 
    reactive({
      temp <- economics_selected_graph()
      temp <- 
        temp[[1]] %>% 
        filter(Year >= input$economics_year_slider[1], Year <= input$economics_year_slider[2])
      
      if(input$economics_selection == economic_factor_choice[1]){
        temp %>% 
            hchart("line", hcaes(`Year`, `Real GDP`)) %>%
            hc_title(text = "Portuguese Real GDP") %>%
            lnplot_custom(.) %>%
            hc_yAxis(title=list(text="<b>Real GDP [int $]</b>", align="middle"),
                     labels = list(
                                formatter = JS("function() {
                                    if (this.value >= 0) {
                                      return (this.value/1000) + ' bil';
                                    }
                                  }")
                                )
                    ) %>%
            hc_xAxis(title=list(text="<b>Year [Quarters per Year]</b>", align="middle")) %>%
            hc_tooltip(.,
                 headerFormat = "",
                 pointFormat =
                   "<b>{point.name}</b><br>Real GDP: {point.y:,.0f} bil int $"
            ) %>%
            hc_caption(
              text = "<b>Real GDP</b> = value expressing the amount of goods and services that are produced and sold through markets, at a given time, adjusted for constant prices using a price deflator.",
              style = list(marginTop = "0px")
            )
      } else if(input$economics_selection == economic_factor_choice[2]){
        temp %>% 
            hchart("line", hcaes(`Year`, `Real GDP Growth Rate`)) %>%
            hc_title(text = "Portuguese Real GDP Growth Rate") %>%
            lnplot_custom(.) %>%
            hc_yAxis(title=list(text="<b>Real GDP Growth Rate [%]</b>", align="middle")) %>%
            hc_xAxis(title=list(text="<b>Year [Quarters per Year]</b>", align="middle")) %>%
            hc_tooltip(.,
                 headerFormat = "",
                 pointFormat =
                   "<b>{point.name}</b><br>Year: {point.x:.f}<br>Real GDP Growth Rate: {point.y:.1f}%"
            ) %>%
            hc_caption(
              text = "<b>Real GDP Growth Rate</b> = growth rate of real GDP calculated year by year.",
              style = list(marginTop = "0px")
            )
      } else if(input$economics_selection == economic_factor_choice[3]){
        temp %>% 
            hchart("line", hcaes(`Year`, `Unemployment Rate`)) %>%
            hc_title(text = "Portuguese Unemployment Rate") %>%
            lnplot_custom(.) %>%
            hc_yAxis(title=list(text="<b>Unemployment Rate [%]</b>", align="middle")) %>%
            hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
            hc_tooltip(.,
                 headerFormat = "",
                 pointFormat =
                   "<b>{point.name}</b><br>Year: {point.x:.f}<br>Unemployment Rate: {point.y:.1f}%") %>%
            hc_caption(
              text = "<b>Unemployment Rate</b> = percentage of the labour force that cannot find occupation within a short time span.",
              style = list(marginTop = "0px")
            )
      } else{
        temp %>% 
            hchart("line", hcaes(`Year`, `Inflation Rate`)) %>%
            hc_title(text = "Portuguese Inflation Rate") %>%
            lnplot_custom(.) %>%
            hc_yAxis(title=list(text="<b>Inflation Rate [%]</b>", align="middle")) %>%
            hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
            hc_tooltip(.,
                 headerFormat = "",
                 pointFormat =
                   "<b>{point.name}</b><br>Year: {point.x:.f}<br>Inflation Rate: {point.y:.1f}%") %>%
            hc_caption(
              text = "<b>Inflation Rate</b> = rate of growth of a price index over time.",
              style = list(marginTop = "0px")
            )
      }
    })
  output$economics_highchart <-
    renderHighchart({
      req(input$economics_selection, input$economics_year_slider)
      economics_filtered_graph()
    })
  }

shinyApp(ui = economics_graph_ui, server = economics_graph_server)
```

<!-- TEXT -->

## General Economic Overview - Text

### **Real GDP and Real GDP Growth Rate**


- General upward trend indicates robust economic growth over time.
- Late 20th Century: Period of rapid expansion, likely driven by:
  - Industrialization.
  - Expansion of the tertiary (service) sector.
  - Integration into the EU.
- Current Era:
  - Stabilisation of growth rate.
- Periods of Stagnation/Contraction:
  - Linked to global external events such as the COVID-19 pandemic.
  
### **Unemployment Rate**

- Stable Trend (pre-2008): Unemployment rate consistently between 5% and 10%.
- Impact of the 2008-2014 Financial Crisis:
  - Unemployment surged to 16%, then returned to the 5%-10% range after the crisis.
- Global Pandemic Impact:
  - Unemployment saw a similar spike during the pandemic but later stabilized.
- Recent Trends (Post-2022):
  - Rapid rebound in labour participation, signaling recovery and growing confidence in the job market.
  - Increase in unemployment during recovery attributed to mismatches between labour demand and supply,         including skill gaps and regional disparities.

### **Inflation Rate**

- 1970s Stagflation:
  - Inflation spiked significantly during the global stagflation event of the 1970s, marked by high inflation   and unemployment.
- Post-1970s Stability:
  - Inflation remained relatively stable between 2% and 3% after the stagflation period, due to stabilizing     economic policies, especially within the Eurozone.
- Inflation between 2020-2022:
  - Influenced by the COVID-19 pandemic and geopolitical tensions (notably the Russo-Ukrainian conflict),       which caused supply chain disruptions and price fluctuations due to scarcity.

<!-- SUPPLY PAGE -->

# Supply

<!-- GRAPHS -->

## Snack Supply - Graphs {data-width="450"}

### 

```{r SUPPLY GRAPHS, fig.height=10}
#CLEAN SUPPLY DATASETS
clean_volume_price <-
  xlsx_dsets[["portugal_retail_volume_price"]] %>% 
  mutate(Category = case_when(
    Category == "Sweet Biscuits, Snack Bars and Fruit Snacks" ~ "Fruit Snacks, Snack Bars and Sweet Biscuits",
    TRUE ~ Category
  )) %>% 
  filter(Category == "Fruit Snacks, Snack Bars and Sweet Biscuits") %>% 
  select(-`Category`, -`Unit`, -`Current Constant P`)%>% 
  pivot_longer(
    cols = c(-Geography,  -`Data Type`), 
    names_to = "Year", # Create a Year column from column names
    values_to = "Value" # Place data values in the Value column
  ) %>% 
  pivot_wider(names_from = `Data Type`, values_from = Value) %>% 
  mutate(`Price/Volume` = `Retail Value RSP` / `Retail Volume (Tonnes)`)

clean_company_shares <-
  xlsx_dsets[["portugal_company_shares"]] %>%
  mutate(Category = case_when(
    Category == "Sweet Biscuits, Snack Bars and Fruit Snacks" ~ "Fruit Snacks, Snack Bars and Sweet Biscuits",
    TRUE ~ Category
  )) %>% 
  filter(Category == "Fruit Snacks, Snack Bars and Sweet Biscuits") %>% 
  select(-Country, -Category, -Unit) %>%
  filter(complete.cases(.)) %>% 
  filter(Company != "Total")

clean_enterprise_n <-
  xlsx_dsets[["portugal_dsets_estat"]] %>% 
  select(Year, `Enterprise Number`) %>% 
  filter(complete.cases(.)) %>% 
  mutate(Year = as.numeric(Year), `Enterprise Number` = as.numeric(`Enterprise Number`))

clean_employee_n <-
  xlsx_dsets[["portugal_dsets_estat"]] %>% 
  select(Year, `Persons Employed`) %>% 
  filter(complete.cases(.))
  
clean_val_added <-
  xlsx_dsets[["portugal_dsets_estat"]] %>% 
  select(Year, `Value Added Millions`) %>% 
  filter(complete.cases(.))

clean_op_surplus <-
  xlsx_dsets[["portugal_dsets_estat"]] %>% 
  select(Year, `Gross Operating Surplus Million Euro`) %>% 
  filter(complete.cases(.))

clean_wages <-
  xlsx_dsets[["portugal_dsets_estat"]] %>% 
  select(Year, `Wages and Salaries`) %>% 
  filter(complete.cases(.))

clean_investment <-
  xlsx_dsets[["portugal_dsets_estat"]] %>% 
  select(Year, `Investment Rate`) %>% 
  filter(complete.cases(.)) %>% 
  mutate(Year = as.numeric(Year), `Investment Rate` = as.numeric(`Investment Rate`))

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#FILTER VECTORS
supply_graph_choice <- c("Volumes w.r.t Sale Revenue", "Active Enterprises", "N° Employees",
                         "Value Added in Production", "Gross Operating Surplus", 
                         "Cost of Labour",  "Investment Rate", "Market Composition")

supply_graph_code <- list(clean_volume_price, clean_enterprise_n, clean_employee_n,
                          clean_val_added, clean_op_surplus, 
                          clean_wages, clean_investment, clean_company_shares)

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC
supply_graph_ui <-
  fluidPage({
    sidebarLayout(
      sidebarPanel(
        width = NULL,
        pickerInput(
          inputId = "supply_selection",
          label = "Supply Graphs",
          choices = supply_graph_choice,
          selected = supply_graph_choice[1],
          options = list(style = "btn-danger"),
          width = "100%"
        ),
        uiOutput(outputId = "supply_plus_filter")
      ),
      
      mainPanel(
        width = NULL,
        highchartOutput(outputId = "supply_highchart")
      ),

      position = c("left", "right")
    )
  })

supply_graph_server <- function(input, output){
  supply_graph_selected <- reactive({
    supply_graph_code[supply_graph_choice %in% input$supply_selection]
  })
  
  output$supply_plus_filter <- renderUI({
    if(input$supply_selection == supply_graph_choice[1]){
        tagList(
          pickerInput(
          inputId = "volumes_country_selection",
            label = "Geographic Scale",
            choices = c("Portugal", "World"),
            selected = "Portugal",
            options = list(style = "btn-danger"),
            width = "100%"
          ),
          chooseSliderSkin("Flat", color = '#db240c'),
          sliderInput(inputId = "supply_volume_price_slider",
                      label = "Year Interval",
                      min = 2010,
                      max = 2024,
                      value = c(2010, 2024),
                      step = 1,
                      sep = "",
                      dragRange = TRUE
                      )
        )
    } else if(input$supply_selection == supply_graph_choice[2]){
        tagList(
          chooseSliderSkin("Flat", color = '#db240c'),
          sliderInput(inputId = "supply_enterprise_slider",
                      label = "Year Interval",
                      min = 2008,
                      max = 2022,
                      value = c(2008, 2022),
                      step = 1,
                      sep = "",
                      dragRange = TRUE
                      )
        )
    } else if(input$supply_selection == supply_graph_choice[3]){
        tagList(
          chooseSliderSkin("Flat", color = '#db240c'),
          sliderInput(inputId = "supply_employee_slider",
                      label = "Year Interval",
                      min = 2008,
                      max = 2022,
                      value = c(2008, 2022),
                      step = 1,
                      sep = "",
                      dragRange = TRUE
                      )
        )
    } else if(input$supply_selection == supply_graph_choice[4]){
        tagList(
          chooseSliderSkin("Flat", color = '#db240c'),
          sliderInput(inputId = "supply_val_added_slider",
                      label = "Year Interval",
                      min = 2008,
                      max = 2022,
                      value = c(2008, 2022),
                      step = 1,
                      sep = "",
                      dragRange = TRUE
                      )
        )
    } else if(input$supply_selection == supply_graph_choice[5]){
        tagList(
          chooseSliderSkin("Flat", color = '#db240c'),
          sliderInput(inputId = "supply_op_surplus_slider",
                      label = "Year Interval",
                      min = 2008,
                      max = 2022,
                      value = c(2008, 2022),
                      step = 1,
                      sep = "",
                      dragRange = TRUE
                      )
        )
    } else if(input$supply_selection == supply_graph_choice[6]){
        tagList(
          chooseSliderSkin("Flat", color = '#db240c'),
          sliderInput(inputId = "supply_wages_slider",
                      label = "Year Interval",
                      min = 2008,
                      max = 2020,
                      value = c(2008, 2020),
                      step = 1,
                      sep = "",
                      dragRange = TRUE
                      )
        )
    } else if(input$supply_selection == supply_graph_choice[7]){
        tagList(
          chooseSliderSkin("Flat", color = '#db240c'),
          sliderInput(inputId = "supply_investment_slider",
                      label = "Year Interval",
                      min = 2008,
                      max = 2020,
                      value = c(2008, 2020),
                      step = 1,
                      sep = "",
                      dragRange = TRUE
                      )
        )
    } else if(input$supply_selection == supply_graph_choice[8]){
        tagList(
            pickerInput(
            inputId = "company_share_selection",
              label = "Company Share Data",
              choices = c("Average Share Value", "Market Concentration"),
              selected = "Average Share Value",
              options = list(style = "btn-danger"),
              width = "100%"
            ),
            chooseSliderSkin("Flat", color = '#db240c'),
            sliderInput(inputId = "supply_company_index_slider",
                      label = "Company Interval - Decreasing Share Value",
                      min = 1,
                      max = 37,
                      value = c(1, 10),
                      step = 1,
                      sep = "",
                      dragRange = TRUE
                      )
        )
    } 
  })
    
    
  supply_filtered_graph <- reactive({
    temp <- supply_graph_selected()
    
    if(input$supply_selection == supply_graph_choice[1]){
      req(input$volumes_country_selection, input$supply_volume_price_slider)
      temp <-
        temp[[1]] %>% 
        filter(Geography == as.character(input$volumes_country_selection)) %>% 
        filter(Year >= input$supply_volume_price_slider[1], Year <= input$supply_volume_price_slider[2])
    } 
    else if(input$supply_selection == supply_graph_choice[2]){
      req(input$supply_enterprise_slider)
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$supply_enterprise_slider[1], Year <= input$supply_enterprise_slider[2])
    }
    else if(input$supply_selection == supply_graph_choice[3]){
      req(input$supply_employee_slider)
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$supply_employee_slider[1], Year <= input$supply_employee_slider[2])
    }
    else if(input$supply_selection == supply_graph_choice[4]){
      req(input$supply_val_added_slider)
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$supply_val_added_slider[1], Year <= input$supply_val_added_slider[2])
    }
    else if(input$supply_selection == supply_graph_choice[5]){
      req(input$supply_op_surplus_slider)
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$supply_op_surplus_slider[1], Year <= input$supply_op_surplus_slider[2])
    } 
    else if(input$supply_selection == supply_graph_choice[6]){
      req(input$supply_wages_slider)
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$supply_wages_slider[1], Year <= input$supply_wages_slider[2])
    }
    else if(input$supply_selection == supply_graph_choice[7]){
      req(input$supply_investment_slider)
      temp <-
        temp[[1]] %>%
        filter(Year >= input$supply_investment_slider[1], Year <= input$supply_investment_slider[2])
    }
    else if(input$supply_selection == supply_graph_choice[8]){
      req(input$company_share_selection)
      if(input$company_share_selection == "Average Share Value"){
        req(input$supply_company_index_slider)  
        temp <-
            temp[[1]] %>% 
            slice(input$supply_company_index_slider[1]:input$supply_company_index_slider[2]) %>% 
            pivot_longer(cols = -Company,
                         names_to = "Year",
                         values_to = "Value"
                         ) %>% 
            filter(Value != "-") %>% 
            mutate(Year = as.numeric(Year), Value = as.numeric(Value)) %>%
            group_by(Year) %>%
            summarize(Mean_Value = mean(Value, na.rm = TRUE))
      } 
      else if(input$company_share_selection == "Market Concentration"){
        req(input$supply_company_index_slider)  
        temp <-
          temp[[1]] %>% 
          slice(input$supply_company_index_slider[1]:input$supply_company_index_slider[2]) %>% 
          pivot_longer(cols = -Company,
                       names_to = "Year",
                       values_to = "Value"
                       ) %>%
          filter(Value != "-") %>%
          mutate(Year = as.numeric(Year), Value = as.numeric(Value)) %>%
          pivot_wider(names_from = Company, values_from = Value)
      } 
    }
  })
    
  output$supply_highchart <- renderHighchart({
    if(input$supply_selection == supply_graph_choice[1]){
      supply_filtered_graph() %>% 
        hchart("line", hcaes(`Retail Volume (Tonnes)`, `Price/Volume`, name= Year)) %>% 
        hc_title(text = "Average Price - Ratio between Sale Revenue and Production Volume") %>%
        lnplot_custom(.) %>%
        hc_yAxis(title=list(
                         text="<b>Average Price [€]</b>", 
                         align="middle"
                       ),
                 labels = list(
                            formatter = JS("function() {
                                if (this.value >= 0) {
                                  return (this.value) + ' €';
                                }
                              }")
                            )
                ) %>%
        hc_xAxis(title=list(text="<b>Snack Retail Volume [000 tonnes]</b>", align="middle"),
                 labels = list(
                            formatter = JS("function() {
                                if (this.value >= 0) {
                                  return (this.value) + 'k';
                                }
                              }")
                            )) %>%
        hc_tooltip(.,
             headerFormat = "",
             pointFormat =
               "<b>{point.name}</b><br>Average Price: {point.y:.1f} €<br>Retail Volume: {point.x:.f}k tonnes<br>Retail Value RSP: {point.Retail Value RSP:.1f} mil €"
             ) %>% 
        hc_caption(
          text = "<b>Retail Sale Revenue/Volumes</b> = relationship expressing the average price of snacks for a given year, total revenue divided by volumes sold at retail level.<br>Eurostat Codes: sbs_na_sca_r2; sbs_sc_ovw",
          style = list(marginTop = "0px")
        )
    } 
    else if(input$supply_selection == supply_graph_choice[2]){
      supply_filtered_graph() %>% 
        hchart("line", hcaes(Year, `Enterprise Number`)) %>% 
        hc_title(text = "Number of Enterprises") %>%
        lnplot_custom(.) %>% 
        hc_yAxis(title=list(
                         text="<b>N° Enterprises</b>", 
                         align="middle"
                       ),
                 labels = list(
                            formatter = JS("function() {
                                if (this.value >= 1000) {
                                  return (this.value/1000) + 'k';
                                } 
                              }")
                            )
                ) %>%
        hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
        hc_tooltip(.,
             headerFormat = "",
             pointFormat =
               "<b>{point.name}</b><br>Year: {point.x:.f}<br>N° Enterprises: {point.y:.f}"
             ) %>% 
        hc_caption(
          text = "<b>Active Enterprises</b> = number of enterprises active in the food manufacture market in a given year.<br>Eurostat Codes: sbs_na_sca_r2; sbs_sc_ovw",
          style = list(marginTop = "0px")
        )
    }
    else if(input$supply_selection == supply_graph_choice[3]){
      supply_filtered_graph() %>% 
        hchart("line", hcaes(Year, `Persons Employed`)) %>% 
        hc_title(text = "Number of Employees") %>%
        lnplot_custom(.) %>% 
        hc_yAxis(title=list(
                         text="<b>N° Employees</b>", 
                         align="middle"
                       ),
                 labels = list(
                            formatter = JS("function() {
                                if (this.value >= 1000) {
                                  return (this.value/1000) + 'k';
                                } 
                              }")
                            )
                ) %>%
        hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
        hc_tooltip(.,
             headerFormat = "",
             pointFormat =
               "<b>{point.name}</b><br>Year: {point.x:.f}<br>N° Employees: {point.y:.f}"
             ) %>% 
        hc_caption(
          text = "<b>N° Employees</b> = number of employees working for enterprises active in the food manufacture market in a given year.<br>Eurostat Codes: sbs_na_sca_r2; sbs_sc_ovw",
          style = list(marginTop = "0px")
        )
    }
    else if(input$supply_selection == supply_graph_choice[4]){
      supply_filtered_graph() %>% 
        hchart("line", hcaes(Year, `Value Added Millions`)) %>% 
        hc_title(text = "Value Added through Production [mil €]") %>%
        lnplot_custom(.) %>% 
        hc_yAxis(title=list(
                         text="<b>Value Added [mil €]</b>", 
                         align="middle"
                       ),
                 labels = list(
                            formatter = JS("function() {
                                if (this.value >= 0) {
                                  return (this.value) + ' mil €';
                                } else if (this.value >= 1000) {
                                  return (this.value/1000) + ' bil €'
                                }
                              }")
                            )
                ) %>%
        hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
        hc_tooltip(.,
             headerFormat = "",
             pointFormat =
               "<b>{point.name}</b><br>Year: {point.x:.f}<br>Value Added: {point.y:.f} mil €"
             ) %>% 
        hc_caption(
          text = "<b>Value Added through Production</b> = increase in value brought by a production process, quantifiable as the difference between its sale price and the cost of its relative production factors, across all enterprises for a given year.<br>Eurostat Codes: sbs_na_sca_r2; sbs_sc_ovw",
          style = list(marginTop = "0px")
        )
    }
    else if(input$supply_selection == supply_graph_choice[5]){
      supply_filtered_graph() %>% 
        hchart("line", hcaes(Year, `Gross Operating Surplus Million Euro`)) %>% 
        hc_title(text = "Gross Operating Surplus [mil €]") %>%
        lnplot_custom(.) %>% 
        hc_yAxis(title=list(
                         text="<b>Gross Operating Surplus [mil €]</b>", 
                         align="middle"
                       ),
                 labels = list(
                            formatter = JS("function() {
                                if (this.value >= 0) {
                                  return (this.value) + ' mil €';
                                } else if (this.value >= 1000) {
                                  return (this.value/1000) + ' bil €'
                                }
                              }")
                            )
                ) %>%
        hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
        hc_tooltip(.,
             headerFormat = "",
             pointFormat =
               "<b>{point.name}</b><br>Year: {point.x:.f}<br>Gross Operating Surplus: {point.y:.f} mil €"
             ) %>% 
        hc_caption(
          text = "<b>Gross Operating Surplus</b> = enterprise profit on goods and services produced considering labour costs, subset of total profits indicating the weight of the labour production factor across enterprises for a given year.<br>Eurostat Codes: sbs_na_sca_r2; sbs_sc_ovw",
          style = list(marginTop = "0px")
        )
    }
    else if(input$supply_selection == supply_graph_choice[6]){
      supply_filtered_graph() %>% 
        hchart("line", hcaes(Year, `Wages and Salaries`)) %>% 
        hc_title(text = "Aggregate Labour Cost [mil €]") %>%
        lnplot_custom(.) %>% 
        hc_yAxis(title=list(
                         text="<b>Aggregate Labour Cost [mil €]</b>", 
                         align="middle"
                       ),
                 labels = list(
                            formatter = JS("function() {
                                if (this.value >= 0) {
                                  return (this.value) + ' mil';
                                } else if (this.value >= 1000) {
                                  return (this.value/1000) + ' bil'
                                }
                              }")
                            )
                ) %>%
        hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
        hc_tooltip(.,
             headerFormat = "",
             pointFormat =
               "<b>{point.name}</b><br>Year: {point.x:.f}<br>Aggregate Labour Cost: {point.y:.f} mil €"
             ) %>% 
        hc_caption(
          text = "<b>Aggregate Labour Cost</b> = cost of labour, aggregate corporate expenditure in wages and salaries in a given year.<br>Eurostat Codes: sbs_na_sca_r2; sbs_sc_ovw",
          style = list(marginTop = "0px")
        )
    }
    else if(input$supply_selection == supply_graph_choice[7]){
      supply_filtered_graph() %>% 
        hchart(., "line", hcaes(x = Year, y = `Investment Rate`)) %>% 
        hc_title(text = "Investment Rate [%]") %>% 
        lnplot_custom(.) %>% 
        hc_yAxis(title=list(text="<b>Investment Rate [%]</b>", align="middle"),
                 labels = list(
                   formatter = JS("function() {
                     if (this.value) {
                         return (this.value) + '%';
                       }
                     }")
                  )
                 ) %>%  
        hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
        hc_tooltip(.,
             headerFormat = "",
             pointFormat =
               "<b>{point.name}</b><br>Year: {point.x:.f}<br>Investment Rate: {point.y:.f}%"
             ) %>%
        hc_caption(
          text = "<b>Investment Rate</b> = ratio between gross tangible investment to net value added, measure of conversion of investments into tangible profit, in the food manufacture sector for a given year.<br>Eurostat Codes: sbs_na_sca_r2; sbs_sc_ovw",
          style = list(marginTop = "0px")
        )
    }
    else if(input$supply_selection == supply_graph_choice[8]){
        if(input$company_share_selection == "Average Share Value"){
          supply_filtered_graph() %>% 
            hchart("line", hcaes(Year, Mean_Value)) %>%
            hc_title(text = "Average Share Value [%]") %>%
            lnplot_custom(.) %>% 
            hc_yAxis(title=list(
                             text="<b>Average Share Value [%]</b>", 
                             align="middle"
                           ),
                     labels = list(
                                formatter = JS("function() {
                                    if (this.value) {
                                      return (this.value) + '%';
                                    } 
                                  }")
                                )
                    ) %>%
            hc_xAxis(title=list(text="<b>Year</b>", align="middle")) %>%
            hc_tooltip(.,
                 headerFormat = "",
                 pointFormat =
                   "<b>{point.name}</b><br>Year: {point.x:.f}<br>Average Share Value: {point.y:.1f}%"
                 ) %>% 
            hc_caption(
              text = "<b>N-Indexed Average Share Value</b> = average value of the first n companies' share, in terms of market size.<br>Share value expressed as percentage of total market value, indicator of relative wealth of n selected companies, in particular when n < 10.<br>Eurostat Codes: sbs_na_sca_r2; sbs_sc_ovw",
              style = list(marginTop = "0px")
            ) 
        } else if (input$company_share_selection == "Market Concentration"){
            temp <- supply_filtered_graph() 
            hc_data <- 
              lapply(
                names(temp)[-1], 
                function(company) {
                  list(name = company, data = as.numeric(temp[[company]]))
                }
              )
            highchart() %>%
              hc_chart(type = "column") %>%
              hc_title(text = "Market Concentration [%]") %>%
              hc_plotOptions(column = list(stacking = "normal")) %>%
              hc_add_series_list(hc_data) %>%
              barplot_custom(.) %>% 
              hc_xAxis(categories = temp$Year) %>%
              hc_yAxis(title = list(text = "<b>Aggregate Share Value [%]</b>", align = "middle"),
                       min = 0,
                       labels = list(
                                 formatter = JS("function() {
                                     if (this.value) {
                                       return (this.value) + '%';
                                     } 
                                   }")
                                )
                       ) %>%
              hc_colors(colorRampPalette(c("#db240c", "#046A38", "#FFE900"))(input$supply_company_index_slider[2])) %>% 
              hc_caption(
                text = "<b>N-Indexed Market Concentration</b> = sum of share value of the first n companies, share value expressed in terms of % of total market value.<br>Indicative of market concentration, the greater the value for small n, the more tendentially monopolistic the market.",
                style = list(marginTop = "0px")
              ) %>% 
            hc_tooltip(valueSuffix = "%")
        }
    }
  })
} 

shinyApp(ui = supply_graph_ui, server = supply_graph_server)
```

<!-- TEXT -->

## Snack Supply - Text

### **Until 2012** 

Demand Contraction (Pre-2012):

- Decline in demand coincided with a fall in active enterprises and in-sector employment.
- Likely linked to the banking crisis during the same period.

Market Adjustments:

- Decline in labour cost, investment success rate and gross operating surplus.
- Implies a decrease of confidence in investment and an increase in other production factor cost, namely raw material cost.
- Decrease in competition resulted in lower prices but growing supply, redistribution of market share.

### **From 2016 to 2018**

- Expansion of supply and demand followed by a sudden supply contraction  (2016), with stable demand inferred from rising prices.
- Measured supply volumes returned to prior year’s levels, challenging clear causal links to observed trends.
  - Labour costs grew steadily since 2013, but their trend does not correlate directly with the supply            contraction.
  - Rising trend in gross operating surplus suggests no significant impact from raw material costs.
  - Number of active enterprises and market concentration among top companies remained stable, offering no        clear explanation.
- Contraction isolated to Portugal, likely tied to national economic policies or food manufacturing regulations.
- A local minimum in investment success indicates reduced confidence in Portugal’s food manufacturing sector.

### **2022 - COVID19 Effect**

- Portuguese supply returned to an upward trend, sustained through the COVID-19 pandemic until 2022.
- 2022 Price Drop:
  - Hypothetically linked to post-lockdown preferences for non-retail foods.
  - More plausibly tied to rising active enterprises and gross operating surplus, reflecting increased            competitiveness in food manufacturing.
- Decline in average share value (% of market value) suggests greater sector competitiveness.
- Potential ties to reduced production factor costs exist but lack post-2022 data for confirmation.
- Top 10 companies maintain ~30% market share; top 4 hold ~23%.
- Quasi-oligarchic market structure indicates stable leadership in the Portuguese food manufacturing sector.

<!-- DEMAND PAGE -->

# Demand

<!-- GRAPHS -->

## Snack Demand - Graphs {data-width="450"}

### 

```{r DEMAND GRAPHS, fig.height=8}
#CLEAN DATASETS
clean_total_market_value <-
  xlsx_dsets[["portugal_retail_volume_price"]] %>% 
  filter(Geography == "Portugal", `Data Type` == "Retail Value RSP") %>% 
  select(-Geography, -`Unit`, -`Current Constant P`, -`Data Type`) %>% 
  pivot_longer(
    cols = -Category,
    names_to = "Year",
    values_to = "Value"
  ) %>% 
  group_by(Year) %>% 
  summarize(tot_market_value = sum(Value))
  

clean_volume_parallel <-
  xlsx_dsets[["portugal_retail_volume_price"]] %>% 
  filter(Geography == "Portugal") %>% 
  select(-Geography, -`Unit`, -`Current Constant P`) %>% 
  pivot_longer(
    cols = c(-Category, -`Data Type`),
    names_to = "Year",
    values_to = "Value"
  ) %>% 
  pivot_wider(
    names_from = `Data Type`,
    values_from = Value
  ) %>% 
  mutate(`Price/Volume` = `Retail Value RSP` / `Retail Volume (Tonnes)`) %>% 
  group_by(Category) %>% 
  mutate(price_2010 = `Price/Volume`[Year == 2010], Indexed_Ratio = `Price/Volume`/price_2010) %>% 
  ungroup()

clean_total_parallel <-
  clean_volume_parallel %>% 
  mutate(Category = case_when(
    Category == "Sweet Biscuits, Snack Bars and Fruit Snacks" ~ "Fruit Snacks, Snack Bars and Sweet Biscuits",
    TRUE ~ Category
  )) %>% 
  filter(Category == "Fruit Snacks, Snack Bars and Sweet Biscuits") %>% 
  mutate(volume_2010 = `Retail Volume (Tonnes)`[Year == 2010], 
         Indexed_Ratio = `Retail Volume (Tonnes)`/volume_2010) %>% 
  group_by(Year) %>%
  summarize(
    Category = "Volume Δ - [Fruit Snacks, Snack Bars, Sweet Biscuits]",  
    Indexed_Ratio = sum(Indexed_Ratio, na.rm = TRUE),
    price_2010 = NA  
  ) %>%
  ungroup()

clean_volume_parallel <-
  bind_rows(clean_volume_parallel, clean_total_parallel)

clean_demand_market_size <-
  xlsx_dsets[["portugal_retail_volume_price"]] %>% 
  filter(`Data Type` == "Retail Value RSP", Geography != "World") %>% 
  select(-Geography, -Unit, -`Data Type`, -`Current Constant P`) %>% 
  mutate(Category = case_when(
    Category == "Sweet Biscuits, Snack Bars and Fruit Snacks" ~ "Fruit Snacks, Snack Bars and Sweet Biscuits",
    TRUE ~ Category
  )) %>% 
  pivot_longer(cols = -Category,
               names_to = "Year",
               values_to = "Value") %>% 
  group_by(Year) %>% 
  mutate(Total = sum(Value, na.rm = TRUE),
         Percentage = (Value/Total)*100) %>% 
  ungroup() %>% 
  select(-Value, -Total) %>% 
  pivot_wider(names_from = Category, values_from = Percentage)

clean_subcategories <-
  xlsx_dsets[["portugal_snack_subcategories"]]%>% 
  select(-Geography, -Datatype, -Unit, -`Current Constant P`) %>% 
  pivot_longer(cols = c(-Category, -Subcategory),
               names_to = "Year",
               values_to = "Value")

clean_gdp_capita_volumes <-
  xlsx_dsets[["portugal_retail_volume_price"]] %>% 
  filter(`Data Type` == "Retail Volume (Tonnes)", Geography != "World") %>% 
  select(-Geography, -`Data Type`, -Unit, -`Current Constant P`) %>% 
  pivot_longer(cols = -Category,
               names_to = "Year",
               values_to = "Value") %>% 
  group_by(Year) %>% 
  summarise(total_vol = sum(Value)) %>% 
  mutate(volume_2010 = total_vol[Year == 2010], 
         i_tot_vol_2010 = total_vol/volume_2010) %>% 
  ungroup() %>% 
  select(-volume_2010)

clean_demand_gdp_capita <-
  dsets[["portugal_gdp_per_capita"]] %>% 
  filter(Year >= 2010 & Entity == "Portugal") %>% 
  select(-Code, -Entity) %>% 
  rename(gdp_capita = GDP.per.capita..PPP..constant.2017.international...) %>% 
  mutate(capita_2010 = gdp_capita[Year == 2010],
         i_gdp_capita_2010 = gdp_capita/capita_2010) %>% 
  select(-capita_2010)

clean_gdp_capita_volumes <-
  merge(clean_gdp_capita_volumes, clean_demand_gdp_capita, by = "Year") 

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#FILTER VECTORS
demand_graph_code <- list(clean_total_market_value, 
                          clean_demand_market_size, clean_subcategories, clean_subcategories,
                          clean_volume_parallel, clean_gdp_capita_volumes)
demand_graph_choice<- c("Total Market Value", "Percentage of Snack Revenue", "Subcategories of Case Study - Values",
                        "Subcategories of Case Study - %", "Volumes Demanded", 
                        "GDP per Capita w.r.t Retail Volumes")

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC
demand_graph_ui <-
  fluidPage({
    sidebarLayout(
      sidebarPanel(
        width = NULL,
        pickerInput(
          inputId = "demand_selection",
          label = "Demand Graphs",
          choices = demand_graph_choice,
          selected = demand_graph_choice[1],
          options = list(style = "btn-danger"),
          width = "100%"
        ),
        uiOutput(outputId = "demand_plus_filter")
        ),

      mainPanel(
        width = NULL,
        highchartOutput(outputId = "demand_highchart")
      ),

      position = c("left", "right")
    )
  })

demand_graph_server <- function(input, output){
  demand_selected <- reactive({
    demand_graph_code[demand_graph_choice %in% input$demand_selection]
  })
  
  output$demand_plus_filter <- renderUI({
    if(input$demand_selection == demand_graph_choice[1]){
      tagList(
        chooseSliderSkin("Flat", color = '#db240c'),
        sliderInput(inputId = "market_val_year_slider",
                    label = "Year Interval",
                    min = 2010,
                    max = 2024,
                    value = c(2010,2024),
                    step = 1,
                    sep = "",
                    dragRange = TRUE
                    )
      )
    }
    else if(input$demand_selection == demand_graph_choice[2]){
      tagList(
        chooseSliderSkin("Flat", color = '#db240c'),
        sliderInput(inputId = "demand_cat_rev_year_slider",
                    label = "Year Interval",
                    min = 2010,
                    max = 2024,
                    value = c(2010,2024),
                    step = 1,
                    sep = "",
                    dragRange = TRUE
                    )
      )
    }
    else if(input$demand_selection == demand_graph_choice[3]){
      tagList(
        chooseSliderSkin("Flat", color = '#db240c'),
        sliderInput(inputId = "demand_subcat_value_year_slider",
                    label = "Year Interval",
                    min = 2010,
                    max = 2024,
                    value = 2010,
                    step = 1,
                    sep = ""
                    ,animate =
                      animationOptions(
                                  interval = 1500,  
                                  loop = FALSE,     
                                  playButton = TRUE
                                  
                      )
                    )
      )
    }
    else if(input$demand_selection == demand_graph_choice[4]){
      tagList(
        chooseSliderSkin("Flat", color = '#db240c'),
        sliderInput(inputId = "demand_subcat_percent_year_slider",
                    label = "Year Interval",
                    min = 2010,
                    max = 2024,
                    value = 2010,
                    step = 1,
                    sep = ""
                    ,animate =
                      animationOptions(
                                  interval = 1500,  
                                  loop = FALSE,     
                                  playButton = TRUE
                                  
                      )
                    )
      )
    }
    else if(input$demand_selection == demand_graph_choice[5]){
      tagList(
        chooseSliderSkin("Flat", color = '#db240c'),
        sliderInput(inputId = "demand_vol_demand_year_slide",
                    label = "Year Interval",
                    min = 2010,
                    max = 2024,
                    value = c(2010,2024),
                    step = 1,
                    sep = "",
                    dragRange = TRUE
                    )
      )
    }
    else if(input$demand_selection == demand_graph_choice[6]){
      tagList(
        chooseSliderSkin("Flat", color = '#db240c'),
        sliderInput(inputId = "gdp_capita_demand_year_slide",
                    label = "Year Interval",
                    min = 2010,
                    max = 2022,
                    value = c(2010,2022),
                    step = 1,
                    sep = "",
                    dragRange = TRUE
                    )
      )
    }
  })
  
  demand_filtered_graph <- reactive({
    temp <- demand_selected()
    
    if(input$demand_selection == demand_graph_choice[1]){
      req(input$market_val_year_slider)
      
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$market_val_year_slider[1], Year <= input$market_val_year_slider[2])
    }
    else if(input$demand_selection == demand_graph_choice[2]){
      req(input$demand_cat_rev_year_slider)
      
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$demand_cat_rev_year_slider[1], Year <= input$demand_cat_rev_year_slider[2])
    }
    
    else if(input$demand_selection == demand_graph_choice[3]){
     req(input$demand_subcat_value_year_slider)
      
      temp <-
        temp[[1]] %>% 
        filter(Year == input$demand_subcat_value_year_slider) %>% 
        filter(Category != Subcategory) %>% 
        select(-Year) %>% 
        pivot_wider(names_from = Subcategory,
                    values_from = Value) %>% 
        mutate_all(~ replace(., is.na(.), 0))
    } 
    else if(input$demand_selection == demand_graph_choice[4]){
       req(input$demand_subcat_percent_year_slider)
      
      temp <-
        temp[[1]] %>% 
        filter(Year == input$demand_subcat_percent_year_slider) %>% 
        filter(Category != Subcategory) %>%
        group_by(Category) %>%
        mutate(cat_total = sum(Value),
               subcat_percent = (Value/cat_total)*100) %>%
        ungroup() %>%
        select(-Year, -Value, -cat_total) %>%
        pivot_wider(names_from = Subcategory,
                    values_from = subcat_percent) %>%
        mutate_all(~ replace(., is.na(.), 0))
    }
    else if(input$demand_selection == demand_graph_choice[5]){
      req(input$demand_vol_demand_year_slide)
      
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$demand_vol_demand_year_slide[1], Year <= input$demand_vol_demand_year_slide[2])
    }
    else if(input$demand_selection == demand_graph_choice[6]){
      req(input$gdp_capita_demand_year_slide)
      
      temp <-
        temp[[1]] %>% 
        filter(Year >= input$gdp_capita_demand_year_slide[1], Year <= input$gdp_capita_demand_year_slide[2])
      
      temp <- 
        list(temp[, c("Year", "total_vol", "i_tot_vol_2010")], 
             temp[, c("Year", "gdp_capita", "i_gdp_capita_2010")])
    }
  })
  
  output$demand_highchart <-
    renderHighchart({
      if(input$demand_selection == demand_graph_choice[1]){
        demand_filtered_graph() %>% 
          hchart("line", hcaes(`Year`, tot_market_value)) %>% 
          hc_title(text = "Snack Market Aggregate Value [mil €]") %>% 
          lnplot_custom(.) %>% 
          hc_yAxis(title = list(text = "<b>Snack Market Aggregate Value<br>[mil €]</b>", align = "middle"),
                   labels = list(formatter = JS("function() {
                                             if (this.value) {
                                               return (this.value) + ' mil €';
                                             } 
                                           }")
                                )
                   ) %>% 
          hc_xAxis(title = list(text = "<b>Year</b>", align = "middle"),
                   labels = list(formatter = JS("function() {
                                             if (this.value) {
                                               return (this.value+2010);
                                             } 
                                           }")
                                )) %>% 
          hc_tooltip(., 
               headerFormat = "",
               pointFormat = 
                 "<b>{point.name}</b><br>Snack Market Aggregate Value: {point.y:.1f} mil €") %>% 
          hc_caption(
            text = "<b>Snack Market Aggregate Value</b> = aggregate retail revenues of all snack categories in Portugal.",
            style = list(marginTop = "0px")
          )
          
      }
      else if(input$demand_selection == demand_graph_choice[2]){
        temp <- demand_filtered_graph()
        
        hc_data <- 
          lapply(
            names(temp)[-1], 
            function(company) {
              list(name = company, data = as.numeric(temp[[company]]))
            }
          )
        
        highchart() %>%
          hc_chart(type = "column") %>%
          hc_title(text = "Snack Categories - Total Retail Revenues Percentage [%]") %>%
          hc_plotOptions(column = list(stacking = "normal")) %>%
          hc_add_series_list(hc_data) %>%
          barplot_custom(.) %>% 
          hc_xAxis(categories = temp$Year) %>%
          hc_yAxis(title = list(text = "<b>Percentage of Total Retail Revenues [%]</b>", align = "middle"),
                   min = 0, max = 100,
                   labels = list(
                             formatter = JS("function() {
                                 if (this.value) {
                                   return (this.value) + '%';
                                 } 
                               }")
                            )
                   ) %>%
          hc_size(height = 400) %>% 
          hc_caption(
            text = "<b>Total Revenues Percentage</b> = individual contributions of other categories of snacks to the total market value.<br>Revenue of single category expressed as percentage of total snack market revenue.",
            style = list(marginTop = "0px")
          ) %>% 
          hc_colors(colorRampPalette(c("#db240c", "#046A38", "#FFE900"))(4)) %>% 
          hc_tooltip(valueSuffix = "%",
                     valueDecimals = 1)
        
      }
      else if(input$demand_selection == demand_graph_choice[3]){
        temp <- demand_filtered_graph()
        
        hc_data <- 
          lapply(
            names(temp)[-1], 
            function(company) {
              list(name = company, data = as.numeric(temp[[company]]))
            }
          )
        highchart() %>%
          hc_chart(type = "column") %>%
          hc_title(text = "Snack Categories - Retail Revenue [mil €]") %>%
          hc_plotOptions(column = list(stacking = "normal")) %>%
          hc_add_series_list(hc_data) %>%
          barplot_custom(.) %>% 
          hc_xAxis(categories = temp$Category) %>%
          hc_yAxis(title = list(text = "<b>Subcategory Retail Revenues [mil €]</b>", align = "middle"),
                   min = 0, max = 350,
                   labels = list(
                             formatter = JS("function() {
                                 if (this.value) {
                                   return (this.value) + ' mil €';
                                 } 
                               }")
                            )
                   ) %>%
          hc_colors(c("#db240c","#FFE900", "#FFB900",colorRampPalette(c("#046A38", "#05fc83"))(5))) %>%
          hc_size(height = 400) %>% 
          hc_caption(
            text = 
              "<b>Retail Revenues</b> = individual contributions of other categories of snacks to the total market value.<br>Variations in consumer preferences shown by variations in snack category consumption.",
            style = list(marginTop = "0px")
          ) %>% 
          hc_tooltip(shared = FALSE,
                     valueSuffix = " mil €",
                     valueDecimals = 1)
      }
      else if(input$demand_selection == demand_graph_choice[4]){
        temp <- demand_filtered_graph()
        
        hc_data <- 
          lapply(
            names(temp)[-1], 
            function(company) {
              list(name = company, data = as.numeric(temp[[company]]))
            }
          )
        highchart() %>%
          hc_chart(type = "column") %>%
          hc_title(text = "Snack Categories - Retail Revenues Percentage [%]") %>%
          hc_plotOptions(column = list(stacking = "normal")) %>%
          hc_add_series_list(hc_data) %>%
          barplot_custom(.) %>% 
          hc_xAxis(categories = temp$Category) %>%
          hc_yAxis(title = list(text = "<b>Percentage of Total Retail Revenues [%]</b>", align = "middle"),
                   min = 0, max = 100,
                   labels = list(
                             formatter = JS("function() {
                                 if (this.value) {
                                   return (this.value) + '%';
                                 } 
                               }")
                            )
                   ) %>%
          hc_colors(c("#db240c","#FFE900", "#FFB900",colorRampPalette(c("#046A38", "#05fc83"))(5))) %>%
          hc_size(height = 400) %>% 
          hc_caption(
            text = "<b>Retail Revenues Percentage</b> = individual contributions of other categories of snacks to the total market value, expressed as percentages.<br>Variations in consumer preferences shown by variations in snack category consumption.",
            style = list(marginTop = "0px")
          ) %>% 
          hc_tooltip(shared = FALSE,
                     valueSuffix = "%",
                     valueDecimals = 1)
      }
      else if(input$demand_selection == demand_graph_choice[5]){
        demand_filtered_graph() %>% 
          group_by(Category) %>% 
          hchart("line", hcaes(x = `Year`, y = `Indexed_Ratio`, group = Category)) %>% 
          hc_title(text = "Snack Categories - Demand Variation",
                   style=list(
                       fontWeight="bold",
                       fontSize= "15px",
                       fontFamily= "Lato"
                      )
                   ) %>%
          hc_yAxis(title = list(text = "<b>Values Indexed 2010</b>", align = "middle"),
                   type="linear",
                   lineWidth=5,
                   lineColor="#ecf0f1"
                   ) %>%
          hc_xAxis(title = list(text = "<b>Year</b>", align = "middle"),
                   lineWidth=5,
                   lineColor="#ecf0f1") %>%
          hc_tooltip(.,
                     shared = TRUE, 
                     crosshairs = TRUE, 
                     valueDecimals = 1
                     ) %>% 
          hc_colors(., c("#002D72", "#046A38", "#FFE900", "#db240c", "#000000")) %>% 
          hc_size(height = 420) %>% 
          hc_caption(
            text = "<b>Indexed Quantity Demanded w.r.t Category Price Variation</b> = average price of every snack category indexed to 2010, overlaid on the volume variation of the category of study, also indexed to 2010 (black line).",
            style = list(marginTop = "0px")
          )
      }
      else if(input$demand_selection == demand_graph_choice[6]){
        temp <- demand_filtered_graph()
        
        highchart() %>% 
          hc_add_series(
            data = temp[[1]],
            type = "line", 
            hcaes(x = Year, y = i_tot_vol_2010),
            name = "Snack Volume Variation",
            tooltip = 
              list(
                pointFormat = "<b>Volume:</b> {point.total_vol:.f} k tonnes<br>"
                ) 
          ) %>% 
          
          hc_add_series(
            data = temp[[2]],
            type = "line",
            hcaes(x = Year, y = i_gdp_capita_2010),
            name = "GDP per Capita Variation",
            tooltip = 
              list(
                pointFormat = "<b>GDP per capita:</b> {point.gdp_capita:,.1f} €/inhab"
              ) 
          ) %>% 
          hc_title(text = "Snack Consumption to Income Variation",
                   style=list(
                       fontWeight="bold",
                       fontSize= "15px",
                       fontFamily= "Lato"
                      )
                   ) %>%
          hc_yAxis(title = list(text = "<b>Values Indexed 2010</b>", align = "middle"),
                   type="linear",
                   lineWidth=5,
                   lineColor="#ecf0f1"
                   ) %>%
          hc_xAxis(title = list(text = "<b>Year</b>", align = "middle"),
                   lineWidth=5,
                   lineColor="#ecf0f1",
                   labels = list(
                                 formatter = JS("function() {
                                     if (this.value) {
                                       return (this.value+2010);
                                     } 
                                   }")
                                ) 
                   ) %>%
          hc_colors(., c("#db240c", "#046A38")) %>% 
          hc_size(height = 400) %>% 
          hc_caption(
            text = "<b>Snack Consumption to Income Variation</b> = comparison between snack consumption and household average available income.",
            style = list(marginTop = "0px")
          ) %>% 
          hc_tooltip(shared = TRUE)
      }
    })
}

shinyApp(ui = demand_graph_ui, server = demand_graph_server)
```

<!-- TEXT -->

## Snack Demand - Text {.tabset .tabset-fade}

### **Demand 1/2** 

#### **Portuguese Market Growth**

- Steady growth from €1,617M in 2010 to a projected €2,210M by 2024.
- **Four Main Categories**:
  - Confectionery **→** Largest share, €109M growth, with a slight decline in share from 31.6% (2010) to 28.1% (2024).
  - Ice Cream **→** Robust growth of €190M since 2010.
  - Savory Snacks **→** Largest proportional growth (+59.6%), increasing share from 18.8% (2010) to 21.9% (2024).
  - Sweet Biscuits, Snack Bars, and Fruit Snacks **→** Decline in share from 25.1% (2015) to 23.5% (2024).
    - Sweet Biscuits:
      - Largest contributor to its category, growing from €334M (2010) to €370M (2024).
      - Plain biscuits are the main driver, rising from €155M (2010) to €183M (2024).
    - Fruit Snacks **→** Steady growth from €24M (2010) to €92M (2024).
    - Snack Bars **→** Modest growth from €48M (2010) to €57M (2024), remaining a smaller segment.
- **Overall Trends**:
  - All four subcategories show consistent growth, with a more balanced distribution of market share compared to 2010.

<div style="height: 40px;"></div>

#### **Average Price w.r.t Snack Volume Output**

- **Consistent Demand Growth** <br>Sweet Snacks show steady demand growth despite gradual price increases, indicating inelastic demand.
- **Consumer Perception** <br>Viewed as essential or habitual, with minimal sensitivity to price changes.
- **Weak Substitution Effects** <br>Demand is unaffected by price drops in other snack categories, suggesting Sweet Snacks meet unique consumer needs (e.g., convenience, perceived health benefits, versatility).
- **Temporary Dip in Demand (2017-18)** <br>Likely due to short-term factors but followed by a rapid recovery, affirming Sweet Snacks' appeal.
- **Consumer Loyalty** <br>Strong consumer loyalty contributes to inelasticity and stable demand.
- **Investment Potential** <br>Sweet Snacks are a promising segment for investment and development due to their market stability.
- **Market Resilience** <br>The combination of inelastic demand and weak substitution effects underlines Sweet Snacks’ robustness as a market category.

### **Demand 2/2**

#### **Income and Consumption**

- **Strong Income-Consumption Correlation** <br>A correlation coefficient of 0.82 (CI: 0.53–0.94) shows a strong relationship between GDP per capita and snack consumption (2010–2023).
- **Snacks as Normal Goods** <br>Consumption increases with rising income, indicating snacks are normal goods.
- **Changing Consumer Behaviours** <br>Snacking has become more integrated into daily life as living standards in Portugal improve.
- **Resilience During Economic Downturns** <br>Snack consumption continued to grow during the 2020 COVID-19 crisis, even as GDP per capita dropped to 2016 levels.
- **Lockdown Influence** <br>Lockdowns reinforced snacking habits, with consumers seeking comfort and convenience while spending more time at home.
- **Deeply Embedded Habits** <br>Snacking habits appear resilient to short-term income fluctuations.
- **Evidence of Market Stability** <br>The consistent upward trajectory of snack consumption highlights the market's robustness and integration into consumer lifestyles.

<div style="height: 15px;"></div>

#### **Shift in Lifestyle**

- **Rising Vegetarianism** <br>Vegetarianism in Portugal increased by over 400% from 2007 to 2017, signaling a significant dietary shift.
- **Veganism on the Rise** <br>Veganism is gaining popularity, particularly among young adults (25–30 years old) and women.
- **Flexitarian Movement** <br>Many identify as flexitarians, reducing meat consumption without strictly adhering to vegetarian or vegan diets.
- **Public Policy Support** <br>Legislation mandates vegan options in public canteens, supporting the demand for plant-based products.
- **Environmental and Ethical Awareness** <br>Awareness of environmental issues and animal welfare, boosted by public figures, drives interest in plant-based diets.
- **Low Overall Adoption** <br>Despite growth, vegetarians and vegans remain a minority compared to omnivores in Portugal.
- **Opportunities for Food Companies** <br>Companies must innovate and expand plant-based offerings to meet growing consumer demand.

<!-- COMPANY PAGE -->

# Vieira de Castro & Filhos Lda.

## Vieira - Text & Stats {data-width="1100"}

### **Information** 

filler

### **Stats**

filler





## Vieira - Images

### **Images**

```{r VIEIRA GRAPH}
#CLEAN DATASETS
clean_vieira_share <-
  xlsx_dsets[["portugal_company_shares"]] %>% 
  filter(Company == "Vieira de Castro Produtos Alimentares SA") %>% 
  select(-Country, -Unit, -Company) %>% 
  pivot_longer(cols = -Category,
               names_to = "Year",
               values_to = "share_val") %>% 
  mutate(share_val = as.numeric(share_val), Year = as.numeric(Year))%>% 
  group_by(Year) %>% 
  mutate(Total = sum(share_val)) %>% 
  ungroup(Year) 


#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC
vieira_graph_ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      width = 4,
      chooseSliderSkin("Flat", color = '#db240c'),
      sliderInput(inputId = "vieira_year_slider",
                  label = "Year Interval",
                  min = 2015,
                  max = 2024,
                  value = 2015,
                  step = 1,
                  sep = "",
                  dragRange = TRUE
                  )
    ),
    mainPanel(
      width = 6,
      highchartOutput("vieira_highchart")
    ),
    position = "left"
  )
)

vieira_graph_server <- function(input, output){
  filtered_vieira <- reactive({
    req(input$vieira_year_slider)

    temp <-
      clean_vieira_share %>%
      filter(Year == input$vieira_year_slider)
  })

  output$vieira_highchart <- renderHighchart({
    temp <- filtered_vieira()
    
    hchart(temp, type = "column", hcaes(x = "Total", y = share_val, color = Category)) %>%
    hc_plotOptions(
      series = list(
        stacking = "normal"  # This ensures the values are stacked
      )
    ) %>% 
    barplot_custom(.)
  })
}

shinyApp(ui = vieira_graph_ui, server = vieira_graph_server)

```

<!-- CONCLUSIONS PAGE -->

# Conclusions

## 

### **Datasets Used**

```{r DATASETS REACTABLE}
#DATASETS
dataset_graph_choice <- c("Consumer Price Index","Daily Caloric Supply","GDP Constant",
                          "GDP Per Capita","HDI","Inflation","Life Expectancy","Migrant Stock",
                          "Overweight Adults","Population","Population Density","Population Growth Rate",
                          "Population Poverty","Real GDP","Unemployment Rate","Company Shares","ESTAT Dataframe",
                          "Retail Volume Price","Snack Subcategories")
  
dataset_graph_code <- append(dsets, xlsx_dsets)

#+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-#
#UI AND SERVER LOGIC

datasets_graph_ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      width = NULL,
      pickerInput(
        inputId = "dataset_selection",
        label = "Raw Datasets",
        choices = dataset_graph_choice,
        selected = dataset_graph_choice[1],
        options = list(style = "btn-danger"),
        width = "100%"
      )
    ),
    mainPanel(
      width = NULL,
      reactableOutput("dataset_reactable")
    ),
    position = "left"
  )
)


dataset_graph_server <- function(input, output){
  selected_dataset <- reactive({
    req(input$dataset_selection)
    
    temp <- dataset_graph_code[dataset_graph_choice %in% input$dataset_selection]
  })
  
  output$dataset_reactable <- renderReactable({
    temp <- selected_dataset()
    temp <- 
      temp[[1]] %>% 
      reactable(
        .,
        paginationType = "jump",
        searchable = TRUE, 
        highlight = TRUE,
        theme = reactableTheme("material"))
  })
}

shinyApp(ui = datasets_graph_ui, server = dataset_graph_server)

```

## Conclusion Tab {.tabset .tabset-fade}

### **Conclusions**

filler

### **Sitography**

filler

sitography:
https://www.geeksforgeeks.org/if-else-if-statement-in-r/
































# Graph Tester

## Graph Tester

###

```{r TESTER}

```

